### 2015-02-02 fig-advection-tracer.tex
put advection contours into their own PDF, include in main PDF

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=4.6in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\centering
\hspace{0.05in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}

\input{../advectTracer-horizontal-snapCol-cubicUpwindCPCFit-divFree/tracer-contour}
\end{document}