### 2015-02-02 fig-advection-error.tex
get advection plots aligned for msc publication

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\hspace{0.15in}\includegraphics[width=2.8in,height=1.32in]{img/schaer-btf-4thorder-error.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour-x-only}
\end{document}