### 2015-02-09 fig-gravityWaves-theta.tex
plot thetaDiff for gravityWaves

\documentclass{article}
\usepackage[paperwidth=5.4in,paperheight=4in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\includegraphics[height=2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}

\includegraphics[height=2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
\end{document}