### 2015-02-10 fig-gravityWaves-theta.tex
add thetaDiff[Zoom] plots

\documentclass{article}
\usepackage[paperwidth=6.4in,paperheight=4.1in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\includegraphics[height=2.2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}
\hspace{0.1in}
\includegraphics[height=4in,angle=270]{../gravityWaves-btf/18000/thetaDiffZoom.eps}
\vspace*{0.1in}

\includegraphics[height=2.2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
\hspace{0.1in}
\includegraphics[height=4in,angle=270]{../gravityWaves-snapCol/18000/thetaDiffZoom.eps}
\vspace*{0.1in}

\includegraphics[height=2.2in,angle=270]{../gravityWaves-diagnostics/thetaDiffLegend.eps}
\hspace{0.2in}
\includegraphics[height=3.9in,angle=270]{../gravityWaves-diagnostics/thetaDiffZoomLegend.eps}
\end{document}