### 2015-02-10 fig-advection-error.tex
add subfigure labels, improve captioning

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}

\setlength{\parindent}{0pt}
\begin{document}
\begin{textblock}{1}(2.5,0.4)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(2.5,8.5)
\normalsize
(b)
\end{textblock}
\hspace{0.35in}\includegraphics[width=2.6in,height=1.32in]{img/schaer-btf-4thorder-error.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
\end{document}