### 2015-03-05 fig-gravityWaves-theta.tex
switch now that plotPatchData spits out PDFs for rawValue plots

\documentclass{article}
\usepackage[paperwidth=6.4in,paperheight=4.1in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{3pt}
\textblockcolour{white}
\begin{textblock}{0.75}(1.2,1.2)
\normalsize
\centering
(a)
\end{textblock}
\begin{textblock}{0.75}(7,1.2)
\normalsize
\centering
(b)
\end{textblock}
\begin{textblock}{0.75}(1.2,8)
\normalsize
\centering
(c)
\end{textblock}
\begin{textblock}{0.75}(7,8)
\normalsize
\centering
(d)
\end{textblock}
\includegraphics[width=2.2in]{../gravityWaves-btf-250m/18000/thetaDiff.pdf}
\hspace{0.1in}
\includegraphics[width=4in]{../gravityWaves-btf-250m/18000/thetaDiffZoom.pdf}
\vspace*{0.1in}

\includegraphics[width=2.2in]{../gravityWaves-snapCol-250m/18000/thetaDiff.pdf}
\hspace{0.1in}
\includegraphics[width=4in]{../gravityWaves-snapCol-250m/18000/thetaDiffZoom.pdf}
\vspace*{0.1in}

\includegraphics[height=2.2in,angle=270]{../gravityWaves-diagnostics/thetaDiffLegend.eps}
\hspace{0.2in}
\includegraphics[height=3.9in,angle=270]{../gravityWaves-diagnostics/thetaDiffZoomLegend.eps}
\end{document}