### 2015-04-27 fig-advection-error.tex
switch advection tests over to AtmosTests

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}

\setlength{\parindent}{0pt}
\begin{document}
\begin{textblock}{1}(2.5,0.4)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(2.5,8.5)
\normalsize
(b)
\end{textblock}
\footnotesize
\vspace*{-2.2em}
\input{../advection-horizontal-linear-btf/error-contour}

\input{../advection-horizontal-cubicUpwind-btf/error-contour}
\end{document}