### 2015-04-27 fig-advection-tracer.tex

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=4.6in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}

\setlength{\parindent}{0pt}
\begin{document}
\begin{textblock}{1}(2.5,0.4)
\normalsize
(a)
\end{textblock}
\footnotesize
\vspace*{-2.2em}
\input{../advection-horizontal-linear-btf/tracer-contour}

\begin{textblock}{1}(2.5,5.5)
\normalsize
(b)
\end{textblock}
\input{../advection-horizontal-cubicUpwind-btf/tracer-contour}

\begin{textblock}{1}(2.5,10.3)
\normalsize
(c)
\end{textblock}
\input{../advection-horizontal-cubicUpwind-cutCell/tracer-contour}
\end{document}