### 2015-05-18 fig-advection-tracer.tex
Plot six panels for advection tracer/error contour

\documentclass{article}
\usepackage[paperwidth=6.4in,paperheight=4.6in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}

\setlength{\parindent}{0pt}
\begin{document}
\input{../advection-horizontal-contours/contours}
\end{document}