### 2015-09-18 src/tf-cutcell-comparison/WellerAppendix.tex
Minor grammar and formatting tweaks to appendix.

\section{Semi-implicit treatment of the Hodge operator\label{appx:SI}}

In order to ensure curl-free pressure gradients, following \citet{weller-shahrokhi2014},
the co-variant momentum component, that is the momentum at the cell
face in the direction between cell centres, is used as the prognostic
variable for velocity:
\begin{equation}
V_{f}=\rho_{f}\mathbf{u}_{f}\cdot\mathbf{d}_{f}
\end{equation}
where $\mathbf{d}_{f}$ is the vector between cell centres and subscript
$f$ means ``at face $f$''. The contravariant momentum component,
that is the flux across faces, is a diagnostic variable:
\begin{equation}
U_{f}=\rho_{f}\mathbf{u}_{f}\cdot\mathbf{S}_{f}
\end{equation}
where $\mathbf{S}_{f}$ is the outward-pointing normal vector to face
$f$ with magnitude equal to the area of the face. If $U$ is the
vector of all values of $U_{f}$ and $V$ is the vector of all values
of $V_{f}$ then we can define the Hodge operator as a matrix that
transforms $V$ to $U$:
\begin{equation}
U=HV.
\end{equation}
For energy conservation, \citet{TC12} showed that the Hodge operator
must be symmetric and positive definite. We define a symmetric $H$
suitable for arbitrary 3D meshes:
\begin{equation}
U_{f}=(\rho\mathbf{u})_{F}\cdot\mathbf{S}_{f}
\end{equation}
where subscript $F$ denotes mid-point interpolation from two surrounding
cell values onto face $f$:
\begin{equation}
(\rho\mathbf{u})_{F}=\frac{1}{2}\sum_{c\in f}(\rho\mathbf{u})_{C}
\end{equation}
where $c\in f$ denotes the two cells sharing face $f$. $(\rho\mathbf{u})_{C}$
is the consistent cell centre reconstruction of $\rho\mathbf{u}$
from surrounding values of $V_{f}$:
\[
(\rho\mathbf{u})_{C}=\left(\sum_{f^{\prime}\in c}\mathbf{d}_{f^{\prime}}\otimes\mathbf{d}_{f^{\prime}}^{T}\right)^{-1}\sum_{f^{\prime}\in c}\mathbf{d}_{f^{\prime}}V_{f^{\prime}}
\]
where $\mathbf{d}_{f^{\prime}}\otimes\mathbf{d}_{f^{\prime}}^{T}$
is a $3\times3$ tensor and so the inversion of the tensor sum is
a local operation which can be calculated once for every cell in the
grid before time-stepping begins. The $H$ implied by this reconstruction
of $U$ is likely to be positive definite for meshes with sufficiently
low non-orthogonality, although this has not been proved.

The semi-implicit technique involves combining the momentum \ref{eq:momentum},
continuity \ref{eq:cont} and $\theta$ \ref{eq:theta} equations
and the equation of state \ref{eq:state} to form a Helmholtz equation
to be solved implicitly, as described by \citet{weller-shahrokhi2014}.
The semi-implicit solution technique with a Hodge operator can be
defined by considering only a discretised form of the continuity equation:
\begin{equation}
\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(HV)^{(n+1)}\right\} =0.\label{eq:contD}
\end{equation}
The divergence is discretised using Gauss' divergence theorem so that:
\begin{equation}
\nabla\cdot\left(HV\right)=\frac{1}{\mathcal{V}_{c}}\sum_{f\in c}n_{f}(HV)_{f}\label{eq:GaussDiv}
\end{equation}
where $\mathcal{V}_{c}$ is the volume of cell $c$, $f\in c$ denotes
the faces of cell $c$, and $n_{f}=1$ if $\mathbf{d}_{f}$ points
outwards from the cell and $n_{f}=$-1 otherwise. Equation \ref{eq:GaussDiv}
is now a sum over a sum since $(HV)_{f}$ is one element of a matrix-vector
multiply. In order to simplify the construction of the matrix for
the Helmholtz problem, only the diagonal terms of $HV$ are treated
implicitly. Therefore, $H$ is separated into a diagonal and off-diagonal
matrix:
\begin{equation}
H=H_{d}+H_{off}.
\end{equation}
Equation \ref{eq:contD} can now be approximated by:
\begin{equation}
\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(H_{d}V)^{(n+1)}+\nabla\cdot(H_{off}V)^{\ell}\right\} =0\label{eq:contDd}
\end{equation}
where superscript $\ell$ denotes lagged values taken from a previous
iteration or from a previous stage of a Runge-Kutta scheme. This was
the approach taken by \citet{weller-shahrokhi2014}. However, the
numerical solution of equation \ref{eq:contDd} turns out to be unstable
when using a large time-step on highly non-orthogonal meshes associated
with terrain following layers over steep orography. Improved stability
and energy conservation can be achieved by splitting $H$ into a diagonal
component which would be correct on an orthogonal grid and a non-orthogonal
correction:
\begin{equation}
H=H_{c}+H_{corr}
\end{equation}
where the diagonal matrix $H_{c}=|\mathbf{S}_{f}|/|\mathbf{d}_{f}|$
and the non-orthogonal correction is $H_{corr}=H-H_{c}$. The orthogonal
part, $H_{c}$, can be treated implicitly in the Helmholtz equation:
\begin{equation}
\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(H_{c}V)^{(n+1)}+\nabla\cdot(H_{corr}V)^{\ell}\right\} =0.\label{eq:contDc}
\end{equation}
This form is used for the solutions of the Euler equations in this
paper and is stable, with good energy conservation for all of the
tests presented.