### 2015-10-02 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
add sampleLine plot

\documentclass{article}
\usepackage{showframe}
\usepackage[paperwidth=6.4in,paperheight=8.3in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\centering
\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
\vspace*{0.1in}
\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
\vspace*{0.1in}
\includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
\vspace*{0.1in}
\includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
\vspace*{0.1in}
\includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
\end{document}