### 2015-10-20 src/tf-cutcell-comparison/fig-thermalAdvection.tex
split thermalAdvection plot into separate figure

\documentclass{article}
\usepackage{showframe}
\usepackage[paperwidth=3.1in,paperheight=3.2in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{3pt}
\textblockcolour{white}
\begin{textblock}{0.75}(2.2,0.4)
\normalsize
\centering
(a)
\end{textblock}
\begin{textblock}{0.75}(2.2,8.4)
\normalsize
\centering
(b)
\end{textblock}
\centering
\includegraphics[width=3in]{../thermalAdvection-btf-250m-50dz/18000/tracerDiffMountain.pdf} \\
\vspace*{0.1in}
\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf}
\end{document}