### 2015-10-21 src/tf-cutcell-comparison/fig-gravityWaves-meshes.tex
plot gravity waves cut cell meshes

\documentclass{article}
\usepackage[paperwidth=6.3in,paperheight=2.1in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{2pt}
\textblockcolour{white}
\begin{textblock}{0.5}(2.7,13.8)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(8,13.8)
\normalsize
(b)
\end{textblock}
\begin{textblock}{1}(13.3,13.8)
\normalsize
(c)
\end{textblock}
\includegraphics{../gravityWaves-mesh-cutCell-250m-300dz/constant/mesh.pdf}
\includegraphics{../gravityWaves-mesh-cutCell-250m-250dz/constant/mesh.pdf}
\includegraphics{../gravityWaves-mesh-cutCell-250m-200dz/constant/mesh.pdf}
\end{document}