### 2016-01-12 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex

\documentclass{article}
\usepackage[paperwidth=6.2in,paperheight=2.6in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}
\usepackage{tikz}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{3pt}
\begin{textblock}{0.75}(11.7,0.3)
\begin{tikzpicture}[
  scale=0.25
]
\draw [dashed, ultra thick] (0,0) -- (0,18);
\end{tikzpicture}
\end{textblock}
\centering
\includegraphics[width=6in]{../gravityWaves-btf-250m-50dz/18000/thetaDiffOverview.pdf} \\
\includegraphics[height=5in,angle=270]{thetaDiffOverview_theta_diff.eps}
\end{document}