### 2016-02-02 src/tf-cutcell-comparison/fig-thermalAdvection.tex
refresh thermalAdvection plots

\documentclass{article}
\usepackage[paperwidth=3.1in,paperheight=5in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{3pt}
\textblockcolour{white}
\begin{textblock}{1.5}(13.3,0.3)
\normalsize
\centering
(a)
\end{textblock}
\begin{textblock}{5.5}(9.3,1)
\normalsize
$\ell_2 = \num{1.12e-7}$
\end{textblock}
\begin{textblock}{1.5}(13.3,7.3)
\normalsize
\centering
(b)
\end{textblock}
\begin{textblock}{5.5}(9.3,8)
\normalsize
$\ell_2 = \num{1.28e-6}$
\end{textblock}
\centering
\includegraphics[width=3in]{../thermalAdvection-btf-250m-50dz/18000/tracerDiffMountain.pdf} \\
\vspace*{0.1in}
\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf} \\
\includegraphics[height=3in,angle=270]{tracerDiffMountain_T_diff.eps}
\end{document}