### 2016-02-05 src/tf-cutcell-comparison/fig-gravityWaves-meshes.tex
more corrections to refer to 'potential temperature' rather than 'tracer density', other small fixes and include line numbers in responses

\documentclass{article}
\usepackage[paperwidth=6.3in,paperheight=2.1in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\centering
\TPMargin{2pt}
\textblockcolour{white}
\begin{textblock}{0.5}(2.7,13.8)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(7.7,13.8)
\normalsize
(b)
\end{textblock}
\begin{textblock}{1}(12.8,13.8)
\normalsize
(c)
\end{textblock}
\includegraphics{../gravityWaves-mesh-cutCell-250m-300dz/constant/mesh.pdf}
\includegraphics{../gravityWaves-mesh-cutCell-250m-200dz/constant/mesh.pdf}
\includegraphics{../gravityWaves-mesh-cutCell-250m-150dz/constant/mesh.pdf}
\end{document}